Dewlap color variation in \textit{Anolis sagrei} is maintained among habitats within islands of the West Indies

### 2020-10-28 ms/sections/methods.tex
backup

We sampled 466 male \textit{Anolis sagrei} from seven islands in the Bahamas Archipelago -- Abaco, North Andros, South Andros, South Bimini, Eleuthera, Long Island, Ragged Island -- and two in the Cayman Islands -- Cayman Brac and Little Cayman (Figure \ref{fig:map}). These islands were chosen to span the breadth of the West Indian range of \textit{Anolis sagrei}, because they have highly similar habitat types,  and because the \textit{A. sagrei} on each island group are derived from ancient and distinct colonization events from Cuba (i.e. relatively evolutionarily independent, \citealt{Reynolds2020}). Three habitats were sampled on each island based on characterizations by \citet{Howard1950} and \citet{Schoener1968}. Each habitat is clearly distinguishable by their dominant vegetation type --- xeric coastal scrub (open, relatively dry habitat consisting of low vegetation or isolated trees), primary coppice forest (closed-canopy forest) and mangrove forest (wet coastal habitat with trees growing in brackish water and high light penetration). Sample sizes are given in Table \ref{suptab:counts}. Our sampling design enabled us to test for differences between habitats at a coarse and fine geographical scale. The median distance between two localities within an island was $\sim 11$km, with some islands being sampled at smaller or larger scales (Figure \ref{supfig:map}, Table \ref{suptab:sites}). $80.3$\% of all pairwise distances within islands were less than $50$km. Additionally, there are no major barriers to dispersal (such as mountains or grassland) on any of the islands that we sampled.

@@

All analyses in this study were performed in R 3.6.1 \citep{RCoreTeam2019}.

@@

In addition to within-island PCA, we performed a PCA on pooled data from the whole archipelago. The first four principal components explained 91.3\% of the variance (Table \ref{suptab:pcavariances}). Again PC1 strongly correlated with brightness (Fig. \ref{supfig:brightness_pooled}, Table \ref{suptab:brightness}). PC2 was positively correlated to short wavelengths (ultraviolet to blue) and negatively correlated to long wavelengths (green to red, Fig. \ref{supfig:pooled}B). PC3 was strongly negatively correlated with UV reflectance and positively correlated with blue-green. PC4 was made of a mosaic of wavelengths, correlating positively with blue and red but negatively with ultraviolet and yellow.\\

@@

Because of the aforementioned violations of the MANOVA assumptions, and to reduce the chances of false discovery, we conducted multivariate group comparisons using support vector machines (SVMs), a model-free, powerful nonparametric supervised machine learning technique.\\

>>>>

We sampled 466 male \textit{Anolis sagrei} from seven islands in the Bahamas Archipelago -- Abaco, North Andros, South Andros, South Bimini, Eleuthera, Long Island, Ragged Island -- and two in the Cayman Islands -- Cayman Brac and Little Cayman (Figure \ref{fig:map}). These islands were chosen to span the breadth of the West Indian range of \textit{A. sagrei}, because they have highly similar habitat types,  and because the \textit{A. sagrei} on each island group are derived from ancient and distinct colonization events from Cuba (i.e. relatively evolutionarily independent, \citealt{Reynolds2020}). Three habitats were sampled on each island based on characterizations by \citet{Howard1950} and \citet{Schoener1968}. Each habitat is clearly distinguishable by its dominant vegetation type --- xeric coastal scrub (open, relatively dry habitat consisting of low vegetation or isolated trees), primary coppice forest (closed-canopy forest) and mangrove forest (wet coastal habitat with trees growing in brackish water and high light penetration). Sample sizes are given in Table \ref{suptab:counts}. Our sampling design enabled us to test for differences between habitats at a coarse and fine geographical scale. The median distance between two localities within an island was $\sim 11$km, with some islands being sampled at smaller or larger scales (Figure \ref{supfig:map}, Table \ref{suptab:sites}), and $80.3$\% of all pairwise distances within islands were less than $50$km. Additionally, there are no major barriers to dispersal (such as mountains or grassland) on any of the islands that we sampled.

@@

All analyses in this study were performed in R 3.6.1 \citep{RCoreTeam2019} and the source code can be found at https://github.com/rscherrer/dewlap, presently private. We used our newly developed R package nmgc (Nested Multivariate Group Comparisons; https://github.com/rscherrer/nmgc), to perform all analyses. This package provides a streamlined and powerful interface to perform group comparison analyses in hierarchical and/or nested by combining both classical statistical tools (e.g. MANOVA, ANOVAs, Kruskal-Wallis) and machine learning tools (e.g. LDA, SVMs) into complete analysis pipelines. The package also facilitates PCA-transformations, assumption testing and automatic choice of suitable posthoc tests, along with full cross-validation, binomial testing of accuracy scores and independent replicate trials for machine learning procedures.

@@

In addition to within-island PCA, we performed a PCA on pooled data from all islands. The first four principal components explained 91.3\% of the variance (Table \ref{suptab:pcavariances}). Again PC1 strongly correlated with brightness (Fig. \ref{supfig:brightness_pooled}, Table \ref{suptab:brightness}). PC2 was positively correlated to short wavelengths (ultraviolet to blue) and negatively correlated to long wavelengths (green to red, Fig. \ref{supfig:pooled}B). PC3 was strongly negatively correlated with UV reflectance and positively correlated with blue-green. PC4 was made of a mosaic of wavelengths, correlating positively with blue and red but negatively with ultraviolet and yellow.\\

@@

Because of the aforementioned violations of MANOVA assumptions, and to reduce the chances of false discovery, we conducted multivariate group comparisons using support vector machines (SVMs), a model-free, powerful nonparametric supervised machine learning technique.\\
